---
title: "Homework 3"
subtitle: "SBD2 - Data-driven Visualization and Decision-Making"
author: "Severin Clauss"
date: Dez. 27, 2023
output:
    html_document:
        df_print: paged
        HTML: default
        toc: true
---

```{r include = FALSE, echo = FALSE}
# Install libraries needed
libraries <- c('igraph')

lapply(libraries, function(x) if (!(x %in% installed.packages())) {
  install.packages(x)
})

lapply(libraries, library, quietly = TRUE, character.only = TRUE)

if (basename(getwd()) != "homework3") {
  setwd("./homework3")
}

set.seed(111)
```

# Task description
1. *Select three movies of your choice (you see the corresponding movie titles in uploaded Excel file).*
2. *Download the network data for the movies of your choice and import it to R (the files are igraph objects, so they are ready to be analyzed). You can do this using the following syntax: graph <- readRDS ("PLACE LINK TO THE FILE HERE").*
3. *Compare the three movies with each other: Calculate network metrics and also visualize the networks (play around with the different options two create visually appealing and readable plots). For this, use the techniques/methods we have covered in class. What insights can you derive from your analysis? --> The goal here really is, to find meaning in the data and carefully interpret relevant network metrics (e.g. How do they differ between the movies and what does this say about the type and/or the plot of the movie?, etc).*
4. *Conduct the entire analysis that you do in R markdown.*
5. *Please use headings to structure your R markdown report and don't forget to clearly indicate, which movies you have chosen.*
6*Upload your work as R markdown file AND additionally upload your knitted report as HTML file.*

# Analysis
For the analysis, I have chosen the following movies:
* A Few Good Men `movie1`
* Above the Law `movie2`
* Absolute Power `movie3`

Let us load the data then
```{r}
movie1 <- readRDS('./Movie_networks/19.rds') # 19 A Few Good Men
movie2 <- readRDS('./Movie_networks/21.rds') # 21 Above the Law
movie3 <- readRDS('./Movie_networks/22.rds') # 22 Absolute Power
```

## Network visualization
### Network graph
```{r message=FALSE, echo=FALSE, fig.width = 20, fig.height = 8}
par(mfrow=c(1, 3), mar=c(2,2,2,2))
plot(movie1, layout = layout.fruchterman.reingold, main='movie1')
plot(movie2, layout = layout.fruchterman.reingold, main='movie2')
plot(movie3, layout = layout.fruchterman.reingold, main='movie3')
```
In these plots, we can try to identify the main characters:
* Movie1: kaffee and jo
* Movie2: nico
* Movie3: seth and luther

Also, there is a difference between the number of nodes and paths. At the first glimpse, we could come to the conclusion
that more nodes and edges lead to a more complex story.
In this assumption, the movie 2 seems to be the most complex one.

### Heatmap
Let us also see the connections in a heatmap


```{r message=FALSE, echo=FALSE}
palf <- colorRampPalette(c("gold", "dark orange"))

### movie 1
netm1 <- get.adjacency(movie1, sparse = F)
colnames(netm1) <- V(movie1)$name
rownames(netm1) <- V(movie1)$name

heatmap(netm1, Rowv = NA, Colv = NA, col = palf(100), scale="none", margins=c(10,10), main = 'Movie1')

### movie2
netm2 <- get.adjacency(movie2, sparse = F)
colnames(netm2) <- V(movie2)$name
rownames(netm2) <- V(movie2)$name

heatmap(netm2, Rowv = NA, Colv = NA, col = palf(100), scale="none", margins=c(10,10), main = 'Movie2')

### movie3
netm3 <- get.adjacency(movie3, sparse = F)
colnames(netm3) <- V(movie3)$name
rownames(netm3) <- V(movie3)$name

heatmap(netm3, Rowv = NA, Colv = NA, col = palf(100), scale="none", margins=c(10,10), main = 'Movie3')
```

## Network Descriptive Statistics
Let us have a look at the key metrics for the network. We look at them on two different levels:
* Global
* Local

### Global measures
#### Density
```{r}
edge_density(movie1)
edge_density(movie2)
edge_density(movie3)
```

The density seems similar low for each movie.
This could lead to the assumption that this is a key essence for a good movie plot?
To get a better understanding, we would have to investigate more with more movies.

#### Shortest path and distance
For diameter:
```{r}
diameter(movie1, directed=FALSE, weights=NA)
diameter(movie2, directed=FALSE, weights=NA)
diameter(movie3, directed=FALSE, weights=NA)
```

For average path length:
```{r}
average.path.length(movie1, directed=FALSE)
average.path.length(movie2, directed=FALSE)
average.path.length(movie3, directed=FALSE)
```

### Local measures
#### Communities
```{r message=FALSE, echo=FALSE, fig.width = 20, fig.height = 8}
par(mfrow=c(1, 3), mar=c(2,2,2,2))
comm1 <- cluster_infomap(movie1)
V(movie1)$color <- membership(comm1)
plot(movie1, main = 'Movie1')

comm2 <- cluster_infomap(movie2)
V(movie2)$color <- membership(comm2)
plot(movie2, main = 'Movie2')


comm3 <- cluster_infomap(movie3)
V(movie3)$color <- membership(comm3)
plot(movie3, main = 'Movie3')

cat("Number of communities in movie1:", length(comm1), "\n")
cat("Number of communities in movie2:", length(comm2), "\n")
cat("Number of communities in movie3:", length(comm3), "\n")
```

The communities for each movie support our assumption that movie 2 is the most complex one.

#### K-core decomposition
Let's plot the K-Core for each movie

**Movie 1**
```{r fig.width = 20, fig.height = 8, echo = F}
coreness(movie1) # max 5

V(movie1)$coreness <- coreness(movie1)
par(mfrow=c(2, 3), mar=c(0.1,0.1,1,0.1))
set.seed(777)
fr <- layout_with_fr(movie1)
for (k in 1:5){
  plot(movie1, main=paste0(k, '-core shell'), layout=fr)
}
```

**Movie 2**
```{r fig.width = 20, fig.height = 8, echo = F}
coreness(movie2) # max 7

V(movie2)$coreness <- coreness(movie2)
par(mfrow=c(3, 3), mar=c(0.1,0.1,1,0.1))
set.seed(777); fr <- layout_with_fr(movie2)
for (k in 1:7){
  plot(movie2, main=paste0(k, '-core shell'), layout=fr)
}
```

**Movie 3**
```{r fig.width = 20, fig.height = 8, echo = F}
coreness(movie3) # max 4

V(movie3)$coreness <- coreness(movie3)
par(mfrow=c(2, 2), mar=c(0.1,0.1,1,0.1))
set.seed(777); fr <- layout_with_fr(movie3)
for (k in 1:4){
  plot(movie3, main=paste0(k, '-core shell'), layout=fr)
}
```

In the maximum K-Core layer of each movie, movie 1 has the moste nodes where movie 2 as the least.
Even thought movie 2 seems to be complex, the story line seems to evolve around a small group of people.

#### Cliques

**Movie 1**
```{r}
clique_num(movie1)
largest_cliques(movie1)
```

**Movie 2**
```{r}
clique_num(movie2)
largest_cliques(movie2)
```

**Movie 3**
```{r}
clique_num(movie3)
largest_cliques(movie3)
```

The clique overview supports the conclusion stated above.
